I/O simulation model with Linux page cache, integration and evaluation in SimGrid framework

### 2020-12-03 paper_io_sim.tex
comments on results
comments and edits in method

		In this section, we present our approach to model file read and write with 
		page cache, cache eviction and data flushing mechanisms implemented 
		in the Linux kernel. 
		We also detail the design of our simulator and its implementation. 
		Because the simulator is implemented with SimGrid using WRENCH, we 
		name the simulator as WRENCH-Ext, along with the WRENCH simulator,
		which is created with original WRENCH.
		Finally, we describe experimental scenarios to evaluate our simulation model 
		using a standalone Python prototype, WRENCH and WRENCH-Ext simulator.

@@

            We separate our simulation model in two components, the IO
			Controller and the Memory Manager, which together simulate 
			file read/write mechanisms in Linux (Figure~\ref{fig:interaction}). 
			The Memory Manager simulates I/Os to cache, flushing, periodical flushing, 
			and cache eviction, whereas the IO Controller simulates file reads and writes. 			
			To read or write a chunk of file, a simulated application sends a request 
			to the IOController.
			Based on the memory status in Memory Manager, the IOController 
			orchestrates cache and disk reads/writes, flushing, cache eviction 
			to simulate I/O time.

			\begin{figure}
   				\centering
   				\includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
				   \caption{Overview of the simulated writeback page cache.
				   Application sends file read and file write requests to IOController, 
				   then IOController manages orchestrate flushing, eviction, cache and 
				   disk accesses to fulfill the requests with the MemoryManager object. 
				   Memory and Disk are responsible for simulating I/O time.}
				   \label{fig:interaction}
			\end{figure}

@@

			We leverage existing storage simulation models to simulate the memory,
			which is characterized with storage capacity, read and write bandwidths 
			and device latency. 
			These features are captured in simulation frameworks and applicable 
			to memory simulation \cite{lebre2015}. 
			In addition, leveraging existing disk models allows us 
			to take advantage of bandwidth sharing between concurrent disk I/O to 
			simulate parallel memory accesses.

			\paragraph*{Memory Manager}
			Memory Manager simulates two parallel threads presented in \ref{fig:interaction}.
			The first thread is the main thread, which is responsible for the simulation 
			of data flushing, cache eviction and I/Os to cache. 
			The second thread is a simulated background thread, in which Memory Manager 
			periodically searches and flushes expired dirty data in page cache LRU lists 
			to disks.

			\paragraph*{Page cache LRU lists}
			In Linux kernel, the page cache LRU lists contain file pages. However, 
			due to the large number of file pages, managing lists of pages may 
			require significant computation overhead. 
			Thus, we introduce the concept of a data block as a unit to represent data 
			cached in memory. A data block is a subset of file pages stored in
            page cache that were accessed in the same I/O operation. 
			A data block has information about file name, block size, last access 
			time, a dirty flag that represents whether the data is clean (0) 
			or dirty (1), and the time it is written to cache as dirty data. 
			Blocks can have different sizes and a given file can have multiple 
			data blocks in page cache. In addition, a data block can be split into an 
			arbitrary number of smaller blocks.

@@

			\begin{figure*}		
			\begin{subfigure}{\linewidth}
				\centering
   				\includegraphics[width=\linewidth]{result/multi/figures/multi_local.pdf}
			\end{subfigure}		
			\caption{I/O time of concurrent pipelines with local storage}
			\label{fig:multi_local}
			\end{figure*}

>>>>

		In this section, we present our approach to model file read and write with 
		page cache, cache eviction and data flushing mechanisms implemented 
		in the Linux kernel. We also detail the design and implementation of two simulators:
		(1) a standalone Python prototype used to evaluate the model independently of any other 
		library, and (2) WRENCH-cache, an extension of WRENCH with our page cache model.
		Finally, we describe experimental scenarios to evaluate our simulations in various configurations.

@@

			We separate our simulation model in two components, the IO
			Controller and the Memory Manager, which together simulate 
			file read and write mechanisms in Linux (Figure~\ref{fig:interaction}). 
			To read or write a file chunk, a simulated application sends a request 
			to the IOController. The IOController determines the amount of memory required by the operation,
			interacts with the Memory Manager accordingly, reads or write to cache through the Memory Manager, 
			and finally reads or writes on disk if necessary.
			The Memory Manager simulates flushing, eviction, and I/O to cache.

			\begin{figure}
   				\centering
   				\includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
				   \caption{Overview of the simulated writeback page cache.
				   Applications send file read or write requests to the
				   IOController that orchestrates flushing, eviction, cache
				   and disk accesses with the MemoryManager. Storage
				   devices (memory and disk) simulate file transfers with
				   bandwidth sharing. \tristan{Could you annotate arrows to
				   make them more explicit?}.}
				   \label{fig:interaction}
			\end{figure}

@@

			We leverage existing storage simulation models~\cite{lebre2015} to simulate the memory,
			characterized by its storage capacity, read and write bandwidths,
			and device latency.
			Leveraging existing storage models allows us 
			to take advantage of bandwidth sharing between concurrent I/Os to 
			simulate parallel memory accesses.

			\paragraph*{Memory Manager}
			The Memory Manager simulates  two parallel threads: the first and main one simulates 
			flushing, eviction, and cache I/Os synchronously; the second one is a background thread that  
			periodically searches for expired dirty data in page cache LRU lists 
			and flushes them to disk.

			\paragraph*{Page cache LRU lists}
			In the Linux kernel, the page cache LRU lists contain file pages. However, 
			due to the large number of file pages, managing lists of pages may 
			require significant computation overhead. 
			Thus, we introduce the concept of a data block as a unit to represent data 
			cached in memory. A data block is a subset of file pages stored in
            page cache that were accessed in the same I/O operation. 
			A data block has information about file name, block size, last access 
			time, a dirty flag that represents whether the data is clean (0) 
			or dirty (1), and the time it is written to cache as dirty data. 
			Blocks can have different sizes and a given file can have multiple 
			data blocks in page cache. In addition, a data block can be split into an 
			arbitrary number of smaller blocks.

@@

			\begin{figure*}		
			\begin{subfigure}{\linewidth}
				\centering
   				\includegraphics[width=\linewidth]{result/multi/figures/multi_local.pdf}
			\end{subfigure}		
			\caption{I/O time of concurrent pipelines with local storage. \tristan{revise caption for real execution variability: "real execution" doesnt tell what the gray shade is.}}
			\label{fig:multi_local}
			\end{figure*}
